---
title: "Data Import"
output: html_document
---

```{r}
library(tidyverse)
```

Let's import the datasets

```{r}
litters_df = 
  read_csv("data_import_examples/FAS_litters.csv", na = c("NA", ".", ""))
litters_df = 
  janitor::clean_names(litters_df)

pups_df = 
  read_csv("data_import_examples/FAS_pups.csv", skip = 3, na = c("NA", ".", ""))
pups_df = 
  janitor::clean_names(pups_df)
```

##`select`

`select` selects variables from my dataframe.

select specific variables

```{r}
select(litters_df, group, litter_number, gd0_weight)
```

select a range

```{r}
select(litters_df, group:gd_of_birth) #this keeps all the variables between group and gd_of_birth
```

select by removal

```{r}
select(litters_df, -group) #this keeps everything BUT group
select(litters_df, -(group:gd_of_birth))
```

select by prefix

```{r}
select(litters_df, group, starts_with("gd")) #keeps everything that starts with "gd" (there is also an "ends_with" or "contains")
```

rearrange by selecting

```{r}
select(litters_df, starts_with("gd"), group) #everything that starts with "gd" will appear first before group
select(litters_df, litter_number, everything()) #puts litter_number first, but keeps all your variables

relocate(litters_df, litter_number) #does the same thing, puts litter_number first (just another option)
```

rename by selecting

```{r}
select(litters_df, GROUP = group, everything()) #just renamed that variable group, but kept everything
rename(litters_df, GROUP = group) #does the same thing, just another method
```

what if I only select one variable?

```{r}
select(litters_df, group) #creates a dataframe with just one variable

pull(litters_df, group) #creates a vector of just the values of one variable. You'll like...never do this
```

LEARNING ASSESSMENT: in the pups data, select the columns containing litter number, sex and PD ears
```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter`

use `filter()` to filter out rows.

```{r}
filter(litters_df, gd_of_birth == 20) #filter keeps everything that's true for gd_of_birth = 20, delete other observations

filter(litters_df, gd_of_birth == 20, pups_born_alive ==6) #filter keeps everything that's true for pups_born_alive = 6 AND the above filter
```

look for more than two pups alive

```{r}
filter(litters_df, pups_born_alive > 5) #all observations where pups_born_alive >5 we KEEP
filter(litters_df, pups_born_alive >= 5)
filter(litters_df, pups_born_alive <= 5)
```

look for non-matches

```{r}
filter(litters_df, pups_born_alive !=6) #rows where NOT equal to 6
```

match to a set
```{r}
filter(litters_df, group %in% c("Con7", "Con8")) #filtering to only keep observations in those 2 groups
```

filter out missing rows with `drop_na()`
```{r}
drop_na(litters_df) #drops ALL rows with NA
drop_na(litters_df, gd0_weight) #only drop missing values in the gd0_weight column, other missing values are kept
```

LEARNING ASSESSMENT: keeping only pups with sex=1. Then, only pups with PD walk less than 11 and sex 2

```{r}
filter(pups_df, sex == 1)
filter(pups_df, pd_walk < 11, sex == 2)
```
























