---
title: "Data Import"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
library(haven)
```

Let's import a dataset.

```{r}
litters_df = 
  read_csv("data_import_examples/FAS_litters.csv")
```
*read_csv is better thab read.csv.

Look at your data!!

```{r}
names(litters_df)
```

Update the names in 'litters_df'.

```{r}
litters_df = 
    janitor::clean_names(litters_df)
```

Look at the data! (If I want, in the console I can do 'view(litters_df)' and see the dataset)

```{r}
litters_df
```

Sometimes skimming data is neat (summaries and stuff)
```{r}
skimr::skim(litters_df)
```

When I look at the dataset I see that there are missing values that are sometimes written as 'NA' and sometimes written as '.' --> both indicate missing values. Let's fix that.

#Fix the missingness

```{r}
litters_df=
    read.csv("data_import_examples/FAS_litters.csv", na = c("NA", ".", ""))
```
 If I want I can 'read.csv("data_import_examples/FAS_litters.csv", na = c("NA", ".", ""), skip = 10) to for example skip the first 10 rows.
 
#Import pups data
 
```{r}
pups_df = 
  read_csv(
    "data_import_examples/FAS_pups.csv", 
    na = c("NA", ".", ""), 
    skip = 3)

pups_df = 
  janitor::clean_names(pups_df)
```

^I imported the pups dataframe and skipped the first 3 rows, then I cleaned the names.

#Okay what about excel

csv's are great, but sometimes you get an excel file.

```{r}
mlb_df = 
  read_excel("data_import_examples/mlb11.xlsx")
```

Import LotR word counts.
```{r}
fotr_df = 
  read_excel("data_import_examples/LotR_Words.xlsx", range = "B3:D6")
```
^I imported just a range of data from the excel.

##SAS??

Import the PULSE data
```{r}
pulse_df = 
  read_sas("data_import_examples/public_pulse_data.sas7bdat")

pulse_df = 
  janitor::clean_names(pulse_df)
```

##Why do I hate read.csv so much??

```{r, eval = FALSE}
litters_df_base =
  read.csv("data_import_examples/FAS_litters.csv")
```

(EVAL=FALSE means I won't execute the code in the readme)
read_csv prints better than read.csv

##What about data exporting?

```{r}
write_csv(fotr_df, "data_import_examples/fotr_df.csv")
```

You do write_csv(dataset_name, "where you want to save it")







